\LARGE \bf Multi-Arm Bandit Models for 2D Sample Based Grasp Planning with Uncertainty [v 17]

### 2015-06-13 paper.tex
Camera Ready

\title{\LARGE \bf
Multi-Arm Bandit Models for 2D Sample Based Grasp Planning with Uncertainty }

\author{Michael Laskey$^1$, Jeff Mahler$^1$, Zoe McCarthy$^1$,  Florian T. Pokorny$^3$, Sachin Patil$^1$,\\ Jur van den Berg$^4$,  Danica Kragic$^3$, Pieter Abbeel$^1$, Ken Goldberg$^2$% <-this % stops a space
\thanks{$^1$Department of Electrical Engineering and Computer Sciences; {\small \{mdlaskey, zmccarthy, jmahler, sachinpatil, pabbeel\}@berkeley.edu}}%
\thanks{$^2$Department of Industrial Engineering and Operations Research and Department of Electrical Engineering and Computer Sciences; {\small goldberg@berkeley.edu}}%
\thanks{$^{1-2}$ University of California, Berkeley;  Berkeley, CA 94720, USA}%
\thanks{$^3$Computer Vision and Active Perception Lab, Centre for Autonomous Systems, School of Computer Science and Communication, KTH Royal Institute of Technology, Stockholm, Sweden {\small \{fpokorny, dani\}@kth.se}}%
\thanks{$^4$Google; Amphitheatre Parkway, Mountain View, CA 94043, USA {\small jurvandenberg@gmail.com}}%
}

@@

\begin{abstract}
For applications such as warehouse order fulfillment, robot grasps must be robust to uncertainty arising from sensing, shape, mechanics, and control.  One way to achieve this is to evaluate the performance of candidate grasps by sampling perturbations in shape, pose, and control, computing the probability of force closure for each candidate to identify the grasp with the highest expected quality. Prior work has turned to cloud computing because evaluating the quality of each grasp is computationally demanding. To improve efficiency and extend this work, we consider how Multi-Armed Bandit (MAB) models for optimizing decisions can be applied in this context. We formulate robust grasp planning as a MAB problem and evaluate grasp planning convergence time using 100 object shapes randomly selected from the Brown Vision 2D Lab Dataset from 1000 uniformly distributed candidate grasp angles. We consider the case where shape uncertainty is represented as a Gaussian process implicit surface (GPIS) and there is Gaussian uncertainty in pose, gripper approach angle, and coefficient of friction. 
Uniform allocation and iterative pruning, the non-MAB methods, converge slowly.
In contrast, Thompson Sampling and the Gittins index method, the MAB methods we consider, converged to within 3\% of the optimal grasp approximately 5x faster.
\end{abstract}

@@

    \caption{Number of samples versus the normalized probability of force closure $P_F$ for the best estimated grasp after $t$ samples, $P_F(\Gamma_{\bar{k}, t})$, out of 1000 candidate grasps using uniform allocation, iterative pruning (eliminating candidates that perform poorly on initial samples), and our proposed Multi-Armed Bandit (MAB) algorithms (Gittins indices and Thompson Sampling).
    %We evaluated performance on 100 objects randomly selected from from the Brown Vision 2D Lab Dataset \cite{brown} with 1000 uniformly distributed candidate grasp plans.
    The normalized $P_F$ is the ratio of $P_F(\Gamma_{\bar{k}, t})$ to the highest $P_F$ in the candidate grasp set $P_F(\Gamma^*)$ averaged over 100 independent runs on randomly selected objects from the Brown Vision 2D Dataset \cite{brown}.
    The highest quality grasp was determined by brute force search over all candidate grasps ( a method that is 10x more iterations than all methods shown)~\cite{kehoe2012toward} 
    Uniform allocation and iterative pruning converge to within 3$\%$ of the optimal (the dashed grey line) in approximately 40,000 and 20,000 iterations, respectively.
    In comparison, the MAB methods both converge in approxmately 4,000 iterations. 
%    $P_F(\Gamma_{\bar{k}})$ (ratio of $P_F(\Gamma_{\bar{k}})$ to $P_F(\Gamma^*)$)
%    
%    Normalized grasp quality over time of the returned grasp for 4 sample-based planning methods.  Normalization divides the current quality of grasp, $P_F(\Gamma_{\bar{k}})$ returned by the highest probability of force closure in the candidate grasp set $P_F(\Gamma^*)$. The highest quality grasp was determined by brute force search over all candidate grasps (which required 10x  more time than any of these methods \cite{kehoe2012toward}).  Data is based on 100 objects randomly selected from from the Brown Vision 2D Lab Dataset \cite{brown} with 1000 uniformly distributed candidate grasp plans. Monte Carlo methods such as Uniform sampling and iterative pruning (eliminating candidates that perform poorly on initial samples) eventually converge to within 3$\%$ of the optimal (the dashed grey line), requiring approximately 40,000 and 20,000 iterations.  Two Multi-Arm Bandit methods perform significantly better, converging after only 4000 iterations..
	}
    \label{fig:grasp_quality}
\end{figure}

One common measure of quality is force closure, the ability to resist external forces and torques in arbitrary directions~\cite{li1988task}. To cope with uncertainty, recent work has explored computing the probability of  force closure given uncertainty in pose~\cite{christopoulos2007handling, kim2012physically, weisz2012pose} and object shape~\cite{kehoe2012estimating, mahler2015gp}.
One way to compute the probability of force closure is using Monte-Carlo integration over sample perturbations in the uncertain quantities\cite{christopoulos2007handling, kim2012physically, weisz2012pose, kehoe2012toward}. However, performing Monte-Carlo integration for each candidate grasp in a set is computationally expensive.  Past work has looked at using parallel programming in the cloud to overcome this problem~\cite{kehoe2012estimating,kehoe2012toward,kehoe2015survey}.
In this work, we aim to extend these cloud-based methods by reducing the number of samples needed to converge to a high-quality grasp.

Our main contribution is formulating this problem as  a Multi-Armed Bandit (MAB) and showing it is possible to allocate sampling effort to only grasps that seem promising \cite{barto1998reinforcement, lai1985asymptotically, robbins1985some}. In a standard MAB there are a set of possible options, or `arms'~\cite{barto1998reinforcement}, that each return a numeric reward from a stationary distribution. The goal in a MAB problem is to  select a sequence of options to maximize expected reward. We formulate the problem of ranking a set of candidate grasps according to a quality metric in the presence of uncertainty as a MAB problem. We then treat the MAB algorithm as an anytime algorithm that terminates once a user defined confidence level is met or at a given stopping time.

We study this formulation using probability of force closure~\cite{christopoulos2007handling, weisz2012pose, kehoe2012toward} as a quality metric under uncertainty in pose, shape, gripper approach, and friction coefficient. 
We model shape uncertainty using Gaussian process implicit surfaces (GPISs), a Bayesian representation of shape uncertainty that has been used in various robotic applications~\cite{dragiev2011, hollinger2013}. 
We model uncertainty in pose as a normal distribution around the orientation and 2D position of the object.
We chose to model uncertainty in grasp approach as a normal distribution around the center and angle of the grasp axis for a parallel jaw gripper. We model uncertainty in friction coefficient as a normal distribution around an expected friction coefficient.

We compare the performance of Thompson sampling and Gittins indices, two popular algorithms for solving the MAB problem, against uniform allocation and an adaptive sampling method known as iterative pruning, which iteratively reduces the set of candidate grasps  based on sample mean~\cite{kehoe2012toward}. In the task of finding grasps with high probability of force closure.
We compare on the Brown Vision 2D Dataset, a dataset of 2D planar objects~\cite{brown, christopoulos2007handling}.

Our initial results in simulation show that Thompson Sampling, a MAB algorithm, required approximately $5\times$ fewer samples than iterative pruning to plan a grasp within $3\%$ of the estimated highest probability of force closure in the set of 1000 randomly selected grasps averaged over 100 objects. 
Although our current implementation is local, our methods may be extended to the cloud by solving $M$ MAB problems on disjoint subsets of the arms and aggregating the results~\cite{glazebrook2000index}.
%by dividing the arms into $M$ subset,where $M$ is the number of available cores, and solving each MAB problem separately\cite{glazebrook2000index}.

@@

Most research on grasp planning focuses on finding grasps by maximizing a grasp quality metric. Grasp quality is often measured by the ability to resist external perturbations to the object in wrench space~\cite{ferrari1992, miller2004graspit}. Analytical quality metrics typically assume precisely known object shape, object pose, and locations of contact~\cite{cheong2011output, ciocarlie2009}. Work on grasping under uncertainty has considered uncertainty in the state of a robotic gripper~\cite{goldberg1990bayesian, stulp2011learning} and uncertainty in contact locations with an object~\cite{zheng2005} Furthermore, recent work has studied the effects of uncertainty in object pose and gripper positioning\cite{brook2011collaborative, hsiao2011bayesian}.

Brook, Ciocarlie, and Hsiao~\cite{brook2011collaborative, hsiao2011bayesian} studied a Bayesian framework to evaluate the probability of grasp success given uncertainty in object identity, gripper positioning, and pose by simulating grasps on deterministic mesh and point cloud models.
Weisz et al.~\cite{weisz2012pose} found that grasps ranked by probability of force closure subject to uncertainty in object pose were empirically more successful on a physical robot than grasps planned using deterministic wrench space metrics. 
Similarly, Kim et al.~\cite{kim2012physically} planned grasps using dynamic simulations over perturbations in object pose. They also found that the planned grasps were more successful on a physical robot than those planned with classical wrench space metrics.

Recent work has also studied uncertainty in object shape, motivated by the use of uncertain low-cost sensors and tolerances in part manufacturing.
Christopoulos et al.~\cite{christopoulos2007handling} sampled spline fits for 2-dimensional planar objects and ranked a set of randomly generated grasps by probability of force closure.
Kehoe et al.~\cite{kehoe2012estimating, kehoe2012toward} sampled perturbations in shape for extruded polygonal objects to plan push grasps for parallel-jaw grippers.
Several recent works have also studied using Gaussian process implicit surfaces (GPISs) to represent shape uncertainty motivated by its ability to model spatial noise correlations and to integrate multiple sensing modalities~\cite{dragiev2011, dragiev2013uncertainty, hollinger2013, mahler2015gp}.
Dragiev et al.~\cite{dragiev2011} uses GPIS to actively explore shapes with tactile sensing to find a hand posture that aligned the gripper fingers to an object's surface normals~\cite{dragiev2013uncertainty}.
Mahler et al. used the GPIS representation to find locally optimal anti-podal grasps by framing grasp planning as an optimization problem~\cite{mahler2015gp}.

@@

Some probabilistic grasp quality measures, such as probability of force closure, are computed using Monte-Carlo integration~\cite{christopoulos2007handling, kehoe2012toward, kim2012physically, weisz2012pose}.
Monte-Carlo involves sampling from distributions on uncertainty quantities and averaging the quality over these samples to empirically estimate a probability distribution~\cite{caflisch1998monte}.
However, it can be computationally expensive to sample all proposed grasps to convergence.
To address the computational cost, Kehoe et al.~\cite{kehoe2012estimating} proposed an adaptive sampling procedure called iterative pruning, which periodically discards a subset of the grasps that seem unlikely to be of high probability of force closure. However, the method pruned grasps using only the sample mean, which could discard good grasps.
We propose modeling the problem as a Multi-Armed Bandit, which selects the next grasp to sample based on past observations instead~\cite{barto1998reinforcement, lai1985asymptotically}.

@@

A traditional MAB example is a gambler with $K$ independent one-armed bandits, also known as slot machines.
When an arm is played (or ``pulled'' in the literature), it returns an amount of money from a fixed reward distribution that is unknown to the gambler.
The goal of the gambler is to come up with a method to maximize the sum of average rewards over all pulls.
If the gambler knew the machine with the highest expected reward, the gambler would only pull that arm.
However, since the reward distributions are unknown, a successful gambler needs to trade off exploiting the arms that currently yields the highest expected reward and exploring new arms.
Developing a policy that successfully trades between exploration and exploitation reward has been the focus of extensive research since the problem formulation  ~\cite{bergemann2006bandit},\cite{bubeck2009pure}, \cite{robbins1985some}.

At each time step the MAB algorithm incurs \textit{regret}, the difference between the expected reward of the best arm and that of the selected arm. Bandit algorithms minimize cumulative regret, the sum of regret over the entire sequence of arm choices.
Lai and Robbins~\cite{lai1985asymptotically} showed that the cumulative regret of the optimal solution to the bandit problem is bounded by a logarithmic function of the number of arm pulls.
They presented an algorithm called (Upper Confidence Bound) UCB that obtains this bound asymptotically~\cite{lai1985asymptotically}.
The algorithm maintains a confidence bound on the distribution of reward based on prior observations and pulls the arm with the highest upper confidence bound.
%Many variants of UCB have thus been proposed~\cite{cesa2006prediction}.

\subsection{Bayesian Algorithms for MAB}\label{sec:bandit_algorithm}
We consider Bayesian MAB algorithms that use previous samples to form a belief distribution on the likelihood of the parameters specifying the distribution of each arm \cite{weber1992gittins,agrawal2011analysis}.  Bayesian methods have been shown empirically to outperform UCB~\cite{chapelle2011empirical, bachman2013greedy}. Bayesian algorithms maintain a belief distribution on the grasp quality distributions for each of the candidate grasps to rank. For instance a Bernoulli random variable $p$ can be used to represent a binary grasping metric like force closure. The prior typically placed on a Bernoulli variable is its conjugate prior, the Beta distribution.
Beta distributions are specified by shape parameters $\alpha$ and $\beta$, where ($\alpha >0$ and $\beta >0$).

One benefit of the Beta prior on Bernoulli reward distributions is that updates to the belief distribution after observing rewards from arm pulls can be derived in closed form.
At timestep $t=0$, we pull arm $k$ and observe reward $R_{k,0}$, where $R_{k,0} \in \lbrace 0, 1\rbrace$.
The posterior of the Beta after this observation is $\alpha_{k, 1} = \alpha_{k, 0} + R_{k,0}$  $\beta_{k, 1} = \beta_{k, 0} + 1-R_{k,0}$, where $\alpha_{k,0}$ and $\beta_{k,0}$ are the prior shape parameters for arm $k$ before any samples are evaluated.

Given the current belief $\alpha_{k, t}, \beta_{k, t}$ for an arm $k$ at time $t$, the algorithm can calculate the expected Bernoulli parameter, $\bar{p}_{k,t}$, as follows:
%Thus, we can write the expected probability of force closure as follows

\vspace{-2ex}
\begin{align}\label{eq:shape_sampling}
\bar{p}_{k,t} = \frac{\alpha_{k,t}}{\alpha_{k,t} + \beta_{k,t}}
\end{align}

@@

\subsubsection{The Gittins Index Method} 
One MAB method is to treat the problem as a Markov Decision Process (MDP) and use Markov Decision theory.
Formally, a MDP is defined as a set of states, a set of actions, a set of transition probabilities between states, a reward function, and a discount factor \cite{barto1998reinforcement}.
In the Beta-Bernoulli MAB case, the set of actions is the $K$ arms and the states are the Beta posterior on each arm, or the integer values of $\alpha_{k,t}$ and $\beta_{k,t}$.

Methods such as Value Iteration can compute optimal policies for a discrete MDP with respect to the discount factor $\gamma$ ~\cite{barto1998reinforcement}.
However, the curse of dimensionality effects performance because for $K$ arms, a finite horizon of $T$ and a Beta-Bernoulli distribution on each arm then the state space is exponential in $K$.
A key insight was given by Gittins, who showed that instead of solving the $K$-dimensional MDP one can instead solve $K$ 1-dimensional optimization problems: for each arm $k$ and for each state $x_{k,t} = \lbrace \alpha_{k,t}, \beta_{k,t} \rbrace$ up to a timestep $T$ \cite{weber1992gittins}.

The solution to the optimization problem assigns each state an index $v(x_{k,t})$.
The indices can then be used to form a policy, where at each timestep the agent selects the arm $k_t$ where $k_t= \underset{1 \leq k \leq K}{\mbox{argmax }} v(x_{k,t})$.  The indices can be  computed offline using a variety of methods \cite{weber1992gittins}; we chose to use the restart method proposed by Katehakis et al. \cite{katehakis1987multi} because it can be implemented in a dynamic programming fashion.
We refer the reader to~\cite{gittins2011multi} for a more detailed analysis of the Gittins index method.

@@

Thompson sampling is an alternative to the Gittins index method that isn't prone to such a problem. 
In Thompson sampling, for each arm draw $p_{k,t} \sim \mbox{Beta}(\alpha_{k,t},\beta_{k,t})$ and pull the arm with the highest $p_{k,t}$ drawn.
A reward is then observed, $R_{k,t}$ and the corresponding Beta distribution is updated. All arms are initialized with prior Beta distributions, which is normally Beta($\alpha_{k,0}=1$,  $\beta_{k,0} =1$) $\forall 1 \leq k \leq K$ to reflect a uniform prior on the parameter of the Bernoulli distribution, $p_{k,0}$.The full algorithm is shown in Algorithm 1.

@@

The intuition for Thompson sampling is that the random samples of $p_{k,t}$ allow the method to explore. However as it receives more samples it hones in on promising arms, since the Beta distributions approach delta distributions as number of samples drawn goes towards infinity \cite{goetschalckx2011continuous}. Chapelle et al. demonstrated empirically that Thompson sampling achieved lower cumulative regret than traditional bandit algorithms like UCB for the Beta-Bernoulli case \cite{chapelle2011empirical}. Theoretically, Agrawal et al. recently proved an upper bound on the asymptotic complexity of  cumulative regret for  Thompson sampling that was sub-linear  for $k$ -arms and logarithmic in the case of $2$ arms  \cite{agrawal2011analysis}.

\section{Grasp Planning Problem Definition}
We consider grasping a rigid, planar object from above using parallel-jaw grippers.
We assume that the interaction between the gripper and object is quasi-static~\cite{kehoe2012estimating, kehoe2012toward}.
We consider uncertainty in shape, pose, gripper approach, and friction coefficient. We assume that the distributions on these quantities are given and can be sampled from.

>>>>

\title{\LARGE \bf
Multi-Arm Bandit Models for 2D Sample Based Grasp Planning with Uncertainty [v 17] }

\author{Michael Laskey$^1$, Jeff Mahler$^1$, Zoe McCarthy$^1$,  Florian T. Pokorny$^3$, Sachin Patil$^1$,\\ Jur van den Berg$^4$,  Danica Kragic$^3$, Pieter Abbeel$^1$, Ken Goldberg$^2$% <-this % stops a space
\thanks{$^1$Department of Electrical Engineering and Computer Sciences; {\small \{mdlaskey , jmahler,zmccarthy, sachinpatil, pabbeel\}@berkeley.edu}}%
\thanks{$^2$Department of Industrial Engineering and Operations Research and Department of Electrical Engineering and Computer Sciences; {\small goldberg@berkeley.edu}}%
\thanks{$^{1-2}$ University of California, Berkeley;  Berkeley, CA 94720, USA}%
\thanks{$^3$Computer Vision and Active Perception Lab, Centre for Autonomous Systems, School of Computer Science and Communication, KTH Royal Institute of Technology, Stockholm, Sweden {\small \{fpokorny, dani\}@kth.se}}%
\thanks{$^4$Google; Amphitheatre Parkway, Mountain View, CA 94043, USA {\small jurvandenberg@gmail.com}}%
}

@@

\begin{abstract}
For applications such as warehouse order fulfillment, robot grasps must be robust to uncertainty arising from sensing, mechanics, and control.  One way to achieve robustness is to evaluate the performance of candidate grasps by sampling perturbations in shape, pose, and gripper approach and compute the probability of force closure for each candidate to identify the grasp with the highest expected quality. Prior work has turned to cloud computing because evaluating the quality of each grasp is computationally demanding. To improve efficiency and extend this work, we consider how Multi-Armed Bandit (MAB) models for optimizing decisions can be applied in this context. We formulate robust grasp planning as a MAB problem and evaluate grasp planning convergence time using 100 object shapes randomly selected from the Brown Vision 2D Lab Dataset from 1000 uniformly distributed candidate grasp angles. We consider the case where shape uncertainty is represented as a Gaussian process implicit surface (GPIS) with Gaussian uncertainty in pose, gripper approach angle, and coefficient of friction. Thompson Sampling and the Gittins index method, the MAB methods we consider, converged to within 3\% of the optimal grasp up to 10x faster than Uniform allocation. 
\end{abstract}

@@

    \caption{Number of samples versus the normalized probability of force closure $P_F$ for the best estimated grasp after $t$ samples, $P_F(\Gamma_{\bar{k}, t})$, out of 1000 candidate grasps using uniform allocation, iterative pruning (eliminating candidates that perform poorly on initial samples), and our proposed Multi-Armed Bandit (MAB) algorithms (Gittins indices and Thompson Sampling).
    %We evaluated performance on 100 objects randomly selected from from the Brown Vision 2D Lab Dataset \cite{brown} with 1000 uniformly distributed candidate grasp plans.
    The normalized $P_F$ is the ratio of $P_F(\Gamma_{\bar{k}, t})$ to the highest $P_F$ in the candidate grasp set $P_F(\Gamma^*)$ averaged over 100 independent runs on randomly selected objects from the Brown Vision 2D Dataset \cite{brown}.
    The highest quality grasp was determined by brute force search over all candidate grasps ( a method that is 10x more iterations than all methods shown)~\cite{kehoe2012toward} 
    Uniform allocation and iterative pruning converge to within 3$\%$ of the optimal (the dashed grey line) in approximately 40,000 and 20,000 iterations, respectively.
    In comparison, the MAB methods both converge in approximately 4,000 iterations. 
%    $P_F(\Gamma_{\bar{k}})$ (ratio of $P_F(\Gamma_{\bar{k}})$ to $P_F(\Gamma^*)$)
%    
%    Normalized grasp quality over time of the returned grasp for 4 sample-based planning methods.  Normalization divides the current quality of grasp, $P_F(\Gamma_{\bar{k}})$ returned by the highest probability of force closure in the candidate grasp set $P_F(\Gamma^*)$. The highest quality grasp was determined by brute force search over all candidate grasps (which required 10x  more time than any of these methods \cite{kehoe2012toward}).  Data is based on 100 objects randomly selected from from the Brown Vision 2D Lab Dataset \cite{brown} with 1000 uniformly distributed candidate grasp plans. Monte Carlo methods such as Uniform sampling and iterative pruning (eliminating candidates that perform poorly on initial samples) eventually converge to within 3$\%$ of the optimal (the dashed grey line), requiring approximately 40,000 and 20,000 iterations.  Two Multi-Arm Bandit methods perform significantly better, converging after only 4000 iterations..
	}
    \label{fig:grasp_quality}
\end{figure}

One common measure of quality is force closure, the ability to resist external forces and torques in arbitrary directions~\cite{li1988task}. To cope with uncertainty, recent work has explored computing the probability of  force closure given uncertainty in pose~\cite{christopoulos2007handling, kim2012physically, weisz2012pose} and object shape~\cite{kehoe2012estimating, mahler2015gp}.
One way to compute the probability of force closure is using Monte-Carlo integration over sampled perturbations in the uncertain quantities~\cite{christopoulos2007handling,kehoe2012toward, kim2012physically, weisz2012pose}. However, performing Monte-Carlo integration for each candidate grasp in a set is computationally expensive.  Past work has looked at using parallel programming in the cloud to overcome this problem and proposed a heuristic for adaptive sampling known as iterative pruning~\cite{kehoe2012estimating,kehoe2012toward,kehoe2015survey}.
In this work, we aim to extend these cloud-based methods by reducing the number of samples needed to converge to a high-quality grasp.

Our main contribution is formulating this problem as  a Multi-Armed Bandit (MAB) and showing that it is possible to allocate sampling effort to grasps with highest estimated probability of force closure\cite{barto1998reinforcement, lai1985asymptotically, robbins1985some}. In a standard MAB, there are a set of possible options, or `arms'~\cite{barto1998reinforcement}, that each return a numeric reward from a stationary distribution. The goal in a MAB problem is to  select a sequence of options to maximize expected reward. We formulate the problem of ranking a set of candidate grasps according to a quality metric in the presence of uncertainty as a MAB problem. We then treat the MAB algorithm as an anytime algorithm that terminates once a user-defined confidence level is met or at a given stopping time.

We study this formulation using probability of force closure~\cite{christopoulos2007handling, weisz2012pose, kehoe2012toward} as a quality metric under uncertainty in pose, shape, gripper approach, and friction coefficient. 
We model shape uncertainty using Gaussian process implicit surfaces (GPISs), a Bayesian representation of shape uncertainty that has been used in various robotic applications~\cite{dragiev2011, hollinger2013}. 
We model uncertainty in pose as a normal distribution around the orientation and 2D position of the object.
Uncertainty in grasp approach is modeled as a normal distribution around the center and angle of the grasp axis for a parallel jaw gripper. Uncertainty in friction coefficient is modeled as a normal distribution around an expected friction coefficient.

We compare the performance of Thompson sampling and Gittins indices, two popular algorithms for solving the MAB problem, against uniform allocation and an adaptive sampling method known as iterative pruning, which iteratively reduces the set of candidate grasps  based on sample mean~\cite{kehoe2012toward}. 
Performance in terms of number of samples needed to find the grasp with highest estimated probability of force closure is measured on the Brown Vision 2D Dataset, a dataset of 2D planar objects~\cite{brown, christopoulos2007handling}.

\todo{Should we say 10x over uniform, or up to 10x improvment, or 5x over iterative pruning?}
Our initial results in simulation show that Thompson Sampling, a MAB algorithm, required $5\times$ fewer samples than iterative pruning to plan a grasp within $3\%$ of the estimated highest probability of force closure in the set of 1000 randomly selected grasps averaged over 100 objects. 
Although our current implementation is local, our methods may be extended to the cloud by solving $M$ MAB problems on disjoint subsets of the arms and aggregating the results~\cite{glazebrook2000index}.
%by dividing the arms into $M$ subset,where $M$ is the number of available cores, and solving each MAB problem separately\cite{glazebrook2000index}.

@@

Many works on grasp planning focus on finding grasps by maximizing a grasp quality metric. Grasp quality is often measured by the ability to resist external perturbations to the object in wrench space~\cite{ferrari1992, miller2004graspit}. Analytical quality metrics typically assume precisely known object shape, object pose, and locations of contact~\cite{cheong2011output, ciocarlie2009}. Work on grasping under uncertainty has considered uncertainty in the state of a robotic gripper~\cite{goldberg1990bayesian, stulp2011learning} and uncertainty in contact locations with an object~\cite{zheng2005}. Furthermore, recent work has studied the effects of uncertainty in object pose and gripper positioning \cite{brook2011collaborative, hsiao2011bayesian}.

Brook, Ciocarlie, and Hsiao~\cite{brook2011collaborative, hsiao2011bayesian} studied a Bayesian framework to evaluate the probability of grasp success given uncertainty in object identity, gripper positioning, and pose by simulating grasps on deterministic mesh and point cloud models.
Weisz et al.~\cite{weisz2012pose} found that grasps ranked by probability of force closure subject to uncertainty in object pose were empirically more successful on a physical robot than grasps planned using deterministic wrench space metrics. 
Similarly, Kim et al.~\cite{kim2012physically} planned grasps using dynamic simulations over perturbations in object pose and found that the planned grasps were more successful on a physical robot than those planned with classical wrench space metrics.

Recent work has also studied uncertainty in object shape, motivated by the use of low-cost sensors and tolerances in part manufacturing.
Christopoulos et al.~\cite{christopoulos2007handling} sampled spline fits for 2-dimensional planar objects and ranked a set of randomly generated grasps by probability of force closure.
Kehoe et al.~\cite{kehoe2012estimating, kehoe2012toward} sampled perturbations in shape for extruded polygonal objects to plan push grasps for parallel-jaw grippers.
Several recent works have also studied using Gaussian process implicit surfaces (GPISs) to represent shape uncertainty motivated by its ability to model spatial noise correlations and to integrate multiple sensing modalities~\cite{dragiev2011, dragiev2013uncertainty, hollinger2013, mahler2015gp}.
Dragiev et al.~\cite{dragiev2011} uses GPIS to actively explore shapes with tactile sensing to find a hand posture that aligned the gripper fingers to an object's surface normals~\cite{dragiev2013uncertainty}.
Mahler et al. used the GPIS representation to find locally optimal antipodal grasps which framed grasp planning as an optimization problem~\cite{mahler2015gp}.

@@

Some probabilistic grasp quality measures, such as probability of force closure, are computed using Monte-Carlo integration~\cite{christopoulos2007handling, kehoe2012toward, kim2012physically, weisz2012pose}.
Monte-Carlo involves sampling from distributions on uncertain quantities and averaging the quality over these samples to empirically estimate a probability distribution~\cite{caflisch1998monte}.
However, it can be computationally expensive to sample all proposed grasps to convergence.
To address the computational cost, Kehoe et al.~\cite{kehoe2012estimating} proposed an adaptive sampling procedure called iterative pruning, which periodically discards a subset of the grasps that seem unlikely to be of high probability of force closure. However, the method pruned grasps using only the sample mean, which could discard good grasps.
We propose modeling the problem as a Multi-Armed Bandit, which selects the next grasp to sample based on past observations instead~\cite{barto1998reinforcement, lai1985asymptotically}.

@@

A traditional MAB example is a gambler with $K$ independent one-armed bandits, also known as slot machines.
When an arm is played (or ``pulled'' in the literature), it returns an amount of money from a fixed reward distribution that is unknown to the gambler.
The goal of the gambler is to come up with a method to maximize the average rewards over all pulls.
If the gambler knew the arm with the highest expected reward, the gambler would only pull that arm.
However, since the reward distributions are unknown, a successful gambler needs to trade off exploiting the arm that currently yields the highest expected reward and exploring new arms.
Developing a policy that successfully trades between exploration and exploitation reward has been the focus of extensive research since the problem formulation~\cite{bergemann2006bandit,robbins1985some}.

At each time step, the MAB algorithm incurs \textit{regret}, the difference between the expected reward of the best arm and that of the selected arm. Bandit algorithms minimize cumulative regret, the sum of regret over the entire sequence of arm choices.
Lai and Robbins~\cite{lai1985asymptotically} showed that the cumulative regret of the optimal solution to the bandit problem is bounded by a logarithmic function of the number of arm pulls.
They presented an algorithm called (Upper Confidence Bound) UCB that obtains this bound asymptotically~\cite{lai1985asymptotically}.
The algorithm maintains a confidence bound on the distribution of reward based on prior observations and pulls the arm with the highest upper confidence bound.

In the robotics field, Hsu et al. applied MAB models to improve the performance and reduce computation time of the probabilistic roadmap motion planner by adaptively sampling waypoints \cite{hsu2005hybrid}. Matikainen et al. formulated policy learning as choosing a state machine from a known libraries of state machines. They then used a MAB algorithm to improve the computational speed of finding the best state machine \cite{matikainen2013multi}. 
%Many variants of UCB have thus been proposed~\cite{cesa2006prediction}.

\subsection{Bayesian Algorithms for MAB}\label{sec:bandit_algorithm}
We consider Bayesian MAB algorithms that use previous samples to form a belief distribution on the parameters specifying the distribution of each arm \cite{agrawal2011analysis,weber1992gittins}.  Bayesian methods have been shown empirically to outperform UCB~\cite{chapelle2011empirical, bachman2013greedy}. Bayesian algorithms maintain a belief distribution on the arm payoff for each of the arms. For instance a Bernoulli random variable $p$ can be used to represent a binary grasping metric like force closure. The prior typically placed on a Bernoulli variable is its conjugate prior, the Beta distribution.
Beta distributions are specified by shape parameters $\alpha$ and $\beta$, where ($\alpha >0$ and $\beta >0$).

One benefit of the Beta prior on Bernoulli reward distributions is that updates to the belief distribution after observing rewards from arm pulls can be derived in closed form.
At timestep $t=0$, we pull arm $k$ and observe reward $R_{k,0}$, where $R_{k,0} \in \lbrace 0, 1\rbrace$.
The posterior of the Beta after this observation is $\alpha_{k, 1} = \alpha_{k, 0} + R_{k,0}$,  $\beta_{k, 1} = \beta_{k, 0} + 1-R_{k,0}$, where $\alpha_{k,0}$ and $\beta_{k,0}$ are the prior shape parameters for arm $k$ before any samples are evaluated.

Given the current belief $\alpha_{k, t}, \beta_{k, t}$ for an arm $k$ at time $t$, the expected Bernoulli parameter, $\bar{p}_{k,t}$, is:
%Thus, we can write the expected probability of force closure as follows

\vspace{-2ex}
\begin{align}\label{eq:shape_sampling}
\bar{p}_{k,t} = \frac{\alpha_{k,t}}{\alpha_{k,t} + \beta_{k,t}} = \frac{\mbox{\#Successes} + \alpha_{k,0}}{\mbox{\#Trials}+\alpha_{k,0}+\beta_{k,0}}
\end{align}

All arms are initialized with prior Beta distributions, which is normally Beta($\alpha_{k,0}=1$,  $\beta_{k,0} =1$) for $k\leq K$ to reflect a uniform prior on the parameter of the Bernoulli distribution, $p_{k,0}$.

@@

\subsubsection{The Gittins Index Method} 
One MAB method is to treat the problem as a Markov Decision Process (MDP) and use Markov Decision theory.
Formally, a MDP is defined by sets of states, actions, transition probabilities between states, a reward function, and a discount factor \cite{barto1998reinforcement}.
In the Beta-Bernoulli MAB case, the set of actions consists of $K$ arms and the set of states are the Beta posterior on each arm, or the values of $\alpha_{k,t}$ and $\beta_{k,t}$.

Methods such as Value Iteration can compute optimal policies for a discrete MDP with respect to the discount factor $\gamma$ ~\cite{barto1998reinforcement}.
However, the curse of dimensionality affects performance because for $K$ arms, a finite horizon of $T$ and a Beta-Bernoulli distribution on each arm then the state space is exponential in size with respect to $K$.
Using the fact that pulling an arm only changes the state of the arm pulled, Gittins showed that instead of solving the $K$-dimensional MDP one can instead solve $K$ 1-dimensional optimization problems: for each arm $k$ and for each state $x_{k,t} = \lbrace \alpha_{k,t}, \beta_{k,t} \rbrace$ up to a timestep $T$ \cite{weber1992gittins}.

The solution to the optimization problem assigns each state an index $v(x_{k,t})$, which can be thought of as the expected value for each state. 
The indices can then be used to form a policy, where at each timestep the agent selects the arm $k_t$ where $k_t= \underset{1 \leq k \leq K}{\mbox{argmax }} v(x_{k,t})$.  The indices for the Beta-Bernoulli case are computed offline and can be found in ~\cite{gittins2011multi}.
We refer the reader to~\cite{gittins2011multi} for a more detailed analysis of the Gittins index method.

@@

@@

Thompson sampling is an alternative to the Gittins index method that is not prone to this problem. 
In Thompson sampling, for each arm draw $p_{k,t} \sim \mbox{Beta}(\alpha_{k,t},\beta_{k,t})$ and pull, the arm with the highest $p_{k,t}$ drawn.
A reward, $R_{k,t}$, is then observed and the corresponding Beta distribution is updated.The general assumption is that sampling from the Beta distribution for each arm is significantly computationally cheaper than actually pulling an arm. Sampling from a Beta distribution is computationally inexpensive and implemented  in most science computing libraries \cite{MATLAB:2010}. The full algorithm is shown in Algorithm 1.

The intuition for Thompson sampling is that the random samples of $p_{k,t}$ allow the method to explore. However, as more samples are received the method focuses on promising arms, since the Beta distributions approach delta distributions as number of samples drawn goes towards infinity \cite{goetschalckx2011continuous}. Chapelle et al. demonstrated empirically that Thompson sampling achieved lower cumulative regret than traditional bandit algorithms like UCB for the Beta-Bernoulli case \cite{chapelle2011empirical}. Agrawal et al. recently proved an upper bound on the asymptotic complexity of  cumulative regret for  Thompson sampling that is sub-linear  for $k$ -arms and logarithmic in the case of $2$ arms  \cite{agrawal2011analysis}.
